---
title: "TMDB Analysis"
author: "Ahmed Abdulhamid"
date: "May 9, 2019"
output: 
  html_document: 
    df_print: paged
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Preprocessing
```{r}
library(tidyverse)
library(jsonlite)

tmdb_5000_credits <- read_csv("tmdb_5000_credits.csv", 
                              col_types = cols(movie_id = col_character()))

tmdb_5000_movies <- read_csv("tmdb_5000_movies.csv", 
                             col_types = cols(id = col_character(), 
                                              runtime = col_double()))

tmdb_5000_movies %>% 
  filter(status == "Released") %>% 
  select(movie_id = id, title, original_language, release_date, budget, revenue, runtime, vote_average, vote_count, popularity) %>% 
  left_join(
    tmdb_5000_movies %>% 
      filter(nchar(genres)>2) %>%
      mutate(js_genres=map(genres,fromJSON)) %>%
      select(id, title, js_genres) %>% 
      unnest(js_genres) %>%
      group_by(id) %>%
      summarise(genres = str_c(name,collapse =  "|")) %>% 
      select(movie_id = id, genres)
  ) %>% 
  left_join(
    tmdb_5000_movies %>% 
      filter(nchar(keywords)>2) %>%
      mutate(js_keywords=map(keywords,fromJSON)) %>%
      select(id, title, js_keywords) %>% 
      unnest(js_keywords) %>%
      group_by(id) %>%
      summarise(keywords = str_c(name,collapse =  "|")) %>% 
      select(movie_id = id, keywords)
  ) %>% 
left_join(
    tmdb_5000_movies %>% 
      filter(nchar(production_companies)>2) %>%
      mutate(js_production_companies=map(production_companies,fromJSON)) %>%
      select(id, title, js_production_companies) %>% 
      unnest(js_production_companies) %>%
      group_by(id) %>%
      summarise(production_companies = str_c(name,collapse =  "|")) %>% 
      select(movie_id = id, production_companies)
) %>% 
  left_join(
    tmdb_5000_movies %>% 
      filter(nchar(production_countries)>2) %>%
      mutate(js_production_countries=map(production_countries,fromJSON)) %>%
      select(id, title, js_production_countries) %>% 
      unnest(js_production_countries) %>%
      group_by(id) %>%
      summarise(production_countries = str_c(name,collapse =  "|")) %>% 
      select(movie_id = id, production_countries)
) -> movies_data


tmdb_5000_credits %>%
  select(movie_id) %>% 
  left_join(
    tmdb_5000_credits %>%
      filter(nchar(crew)>2) %>%
      mutate(js_crew=map(crew,fromJSON)) %>%
      select(-crew, -cast) %>% 
      unnest(js_crew) %>% 
      filter(job == "Director") %>% 
      group_by(movie_id) %>% 
      arrange(movie_id, id) %>%
      slice(1) %>% 
      select(movie_id, director = name)
  ) %>% 
  left_join(
tmdb_5000_credits %>%
  filter(nchar(crew)>2) %>%
  mutate(js_crew=map(crew,fromJSON)) %>%
  select(-crew, -cast) %>% 
  unnest(js_crew) %>% 
  filter(job == "Producer") %>% 
  group_by(movie_id) %>% 
  arrange(movie_id, id) %>%
  slice(1) %>% 
  select(movie_id, producer = name)
) %>% 
left_join(
tmdb_5000_credits %>%
  filter(nchar(cast)>2) %>%
  mutate(js_cast=map(cast,fromJSON)) %>%
  select(-crew, -cast) %>% 
  unnest(js_cast) %>%
  arrange(movie_id, order) %>%
  group_by(movie_id) %>% 
  summarise(actor_1 = nth(name, 1),
         actor_2 = nth(name, 2),
         actor_3 = nth(name, 3))
) -> cast_crew



write_csv(movies_data, "tmdb_movies_metadata.csv")
write_csv(cast_crew, "tmdb_cast_crew.csv")

```

### Calling necessary packages.

```{r}
library(dplyr) 
library(tidyverse)
library(ggplot2)
library(ggpubr)


```

## let's begin by loading our datasets and explore them.
```{r}
df_movies = read_csv("tmdb_movies_metadata.csv")
df_cast_crew = read_csv("tmdb_cast_crew.csv")
```


### Now, we will take a glimpse at movies dataset.
```{r}
glimpse(df_movies)
```

### Same with cast_crew dataset.

```{r}
glimpse(df_cast_crew)
```

### Take a subset from movies to geners

```{r}
df_genres <- select(df_movies, movie_id ,title , vote_average, genres)

```


### Doing Some transformations to reach a form that each row has one genre, with multible rows for the movie.


```{r}
df_genres_final <- separate_rows(df_genres,genres, sep = "\\|")
df_genres_final
```

### In general, what is the average vote related to each genre?


```{r}
df_genres_groubed <- df_genres_final %>%   
                      group_by(genres)  %>%
                      filter(!is.na(genres)) %>% 
                      summarise(average_vote = mean(vote_average))

df_genres_groubed 
```

### Trying to make the answer more visual

```{r}
ggplot(data=df_genres_groubed, aes(x=genres, y=average_vote)) +
      geom_bar(stat="identity") +
      coord_flip()
```
### For the keywords, which of them is related to the highest rating?

```{r}
df_keywords <- select(df_movies, vote_average, keywords)
highest_keywords <- df_keywords %>% separate_rows(keywords, sep = "\\|")   %>% 
                group_by(keywords) %>% 
                filter(!is.na(keywords)) %>% 
                summarise(average_vote = mean(vote_average)) %>% 
                top_n(n = 10 , wt = average_vote)
highest_keywords
```

### Is there any relation between the movie budget and the average vote?

```{r}
budget_rating <- select(df_movies, movie_id,budget, vote_average)
## Normalizing the features
budget_rating$budget <- (budget_rating$budget - mean(budget_rating$budget)) / sd(budget_rating$budget)
budget_rating$vote_average <- (budget_rating$vote_average - mean(budget_rating$vote_average)) / sd(budget_rating$vote_average)

budget_rating
```

### Using some visuals

```{r}
ggscatter(budget_rating, x = "vote_average", y = "budget", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Vote Average", ylab = "Budget")
```
### As shown in the figure above there is no realation between the movie budgut and the average vote 
   Note (Values in the figure are not the actual ones )
   
   
   
### What is the relation between each director and the average vote?


```{r}
movies_rate <- select(df_movies, movie_id , vote_average )
director_actor  <- select(df_cast_crew , movie_id , director , actor_1 , actor_2 ,actor_3)

df_movie_rating_crew <- movies_rate %>% inner_join( director_actor , by = "movie_id")
df_movie_rating_crew %>%  group_by(director) %>% 
                          filter(!is.na(director)) %>% 
                          summarise(average_vote = mean(vote_average)) %>% 
                          top_n(10 )

```
### Who are the actors related to the high rating?


```{r}
Top_10_actors  <-   df_movie_rating_crew %>% 
                    unite("actors" , c("actor_1" , "actor_2" , "actor_3") , remove =TRUE , sep = "|" ) %>% 
                    separate_rows(actors , sep = "\\|") %>% 
                          group_by(actors) %>% 
                          filter(!is.na(actors)) %>% 
                          summarise(average_vote = mean(vote_average)) %>% 
                          top_n(10 )

Top_10_actors
```
### Who are the directors with highest revenue in their movies?


```{r}
director_revenue <- select(df_movies, movie_id , revenue ) %>%
                    inner_join( director_actor , by = "movie_id") %>% 
                    group_by(director) %>% 
                    filter(!is.na(director)) %>% 
                    summarise(revenue = sum(revenue)) %>% 
                    top_n(10 )%>% 
                    arrange(desc(revenue))
director_revenue

```

